Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-05-01 manuscript.tex
Andy edits

%include hyperlinks to figures
%\documentclass{article}
\usepackage{graphicx}
\usepackage[%  
    colorlinks=false,
    pdfborder={0 0 0},
    linkcolor=blue
]{hyperref}

@@

\usepackage[scaled=0.85]{beramono}
\usepackage[T1]{fontenc}
\usepackage{changepage}
\usepackage{mathrsfs}

@@

\newcommand{\state}[1]{$\mathcal{S}_{#1}$}

\newcommand*{\rood}[1]{#1}
\newcommand*{\blauw}[1]{#1}
\newcommand*{\groen}[1]{#1}
\newcommand*{\blauwr}[1]{#1}
\newcommand{\Expect}[1]{\mathrm{E}\left[#1\right]}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}

%\newcommand*{\addref}[1]{\textcolor{red}{\{ADD REF: #1\}}}

@@

%\usepackage[draft]{todonotes}   % notes shown
\usepackage[disable]{todonotes} % notes hidden
\usepackage{comment}

\usepackage[normalem]{ulem}
\usepackage{algorithm}
\usepackage{subfig}
\usepackage{algpseudocode}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[]{Applying State-free Reversible VAMPNets and Markov State Models to Learn Dynamics of DNA Oligonucleotides}

@@

\begin{abstract}

Despite rapid advances in DNA nanotechnology and a robust understanding of the associated thermodynamics, the sequence-dependent mechanisms of DNA hybridization are not fully understood. In this work, we investigate these dynamics by performing equilibrium coarse-grained simulations of oligonucleotide sequences with varied G:C placement. We employ State-Free Reversible VAMPnets to directly learn the slowest dynamical modes of each sequence and optimize Markov State Model construction. Furthermore, we perform elevated temperature simulations to recapitulate temperature-jump IR and FTIR data collected on the oligonucleotides. For repetitive sequences, we find a spectrum of slow dynamics associated with out-of-register base pairing and kinetically relevant transitions between these states. In contrast, G:C pairs near the center of the duplex induce more pronounced fraying dynamics through which hybridization and dissociation are facilitated. In both cases, these mechanisms deviate from an “all-or-nothing” hybridization model. Our computational predictions show agreement with experiments and provide new fundamental understanding of the sequence-dependent kinetics and mechanisms of DNA hybridization.

% revise "excellent accord" above?

\noindent \url{https://www.overleaf.com/project/5e9e5110c524b8000192c548}

\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Over the last couple of decades, DNA has proven to be much more than a vessel for genetic information. From sensing to computing to directed self-assembly, the programmable and predictable nature of DNA has unlocked numerous unforeseen nanotechnology applications.\citep{Seeman2017DNANanotechnology, Adleman1994MolecularProblems, Rothemund2006FoldingPatterns, Gu2010ALine} Recently, single molecule localization techniques have exploited the rapid and transient binding of short DNA oligomers in order to achieve super-resolution microscopy and optical multiplexing. \cite{Schnitzbauer2017Super-resolutionDNA-PAINT, Shah2019, Strauss2020UpDNA-PAINT} Indeed, DNA dynamical behavior has become increasingly important in developing these technologies and understanding fundamental biological processes such as transcription and gene regulation.\citep{Deluca2020DynamicDevices, Cordes2010SensingTransfer, Naimark2020DNADevelopment} A thorough understanding of the hybridization process -- the formation of a DNA duplex from two single strands -- and its associated dynamics is integral to further advancements in these fields. The effect of sequence on hybridization thermodynamics has been rigorously explored, and nearest-neighbor calculations can account for mismatched pairs, dangling ends, and other non-native bonding effects.\citep{SantaLucia1998AThermodynamics, Santalucia2004TM} Secondary DNA structures such as hairpins and G-quadruplexes have also been studied in depth and leveraged for nanotechnology applications.\citep{Tsukanov2013DetailedOrigami, Mosayebi2014TheFormation, Mergny2019DNANanotechnology} Although many experimental and computational studies have investigated DNA dynamical phenomena from the picosecond to millisecond range, the sequence-dependent mechanisms of hybridization and dissociation dynamics are not fully understood.\citep{Yin2011KineticsHybridization, Xiao2019, Hinckley2014Coarse-grainedEffects, Sanstead2016, Porschke1971CooperativeTransition} Moreover, it is unclear the extent to which these processes evolve in an "all-or-nothing" fashion or if long-lived metastable states facilitate the transition.\citep{Araque2016LatticeCooperativity, Sikora2013ModelingIntermediates}

Our understanding of hybridization dynamics has been built from decades of experiments -- such as temperature-jump, salt-jump, pH-jump, and other perturbative methods -- that rapidly stimulate DNA and monitor relaxation to a new equilibrium.\citep{Morrison1993SensitiveSolution, Wetmur1968KineticsDNA, Craig1971RelaxationOligonucleotides, Porschke1973ThermodynamicsPairs, Williams1989LaserDGCATGC, Narayanan2012ExploringMixing, Chen2007InfluenceHybridization, Sanstead2018DirectDehybridization} More recently, single molecule diffusion and tethered multifluorophore assays have facilitated equilibrium analysis, however these results can be hampered by slow data collection rates and fluorescent tags effects on strand dynamics, particularly for shorter oligomers.\citep{Liu20173DSolution,  Schickinger2018TetheredHelices, Chen2008Base-by-baseSpectroscopy, Dupuis2013Single-moleculeHelices} Given that dynamic insights from these experiments are limited, several coarse-grained molecular dynamics (MD) models have been employed to provide molecular level detail.\citep{Romano2013DNADependence, Hinckley2013AnHybridization, Maciejczyk2014DNAModel, Markegard2015, Dans2016MultiscaleDNA} Although these models provide experimentally verified speed-ups compared to all-atom simulations, the long timescales on which DNA hybridization and dissociation events occur can still make these processes difficult to sample via direct simulation techniques.\citep{Phys2014} Instead, many previous studies of DNA hybridization have employed accelerated sampling methods such as metadynamics, umbrella sampling,  transition path sampling, and forward flux sampling.\citep{Hinckley2013AnHybridization, Schmitt2013ExploringSurface, Sambriski2009, Sambriski2009SequencePathways, Hoefert2011MolecularOligonucleotides,Romano2013DNADependence, Hinckley2014Coarse-grainedEffects} Simplified lattice models can expedite computation time and yield useful thermodynamics and transition state information, however some configurations may not be accessible and kinetics are less informative.\citep{Araque2016LatticeCooperativity, Phys2019} Other computational works use elevated temperature or denaturing solvent concentrations to induce one-way dissociation events.\citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding} Taken together, most experimental and computational works have studied the hybridization process as a relaxation to a chosen equilibrium or in a biased fashion.

Recent studies have coupled experimental techniques with machine learning (ML) and  MD simulations to investigate and predict sequence-dependent kinetics.\citep{Schickinger2018TetheredHelices, Zhang2018PredictingSequence} Where these studies focus on association and dissociation kinetics alone, we seek to broaden our analysis into higher-order dynamical processes and metastable intermediates. For example, stable out-of-register or "shifted" base pairing in repetitive sequences have been documented in previous computational studies.\citep{Phys2014, Maciejczyk2014DNAModel, Araque2016LatticeCooperativity, Xiao2019} The extent to which these states are kinetically relevant is difficult to determine experimentally and appears highly sequence-dependent. Frayed structures and dynamics have also been investigated in numerous computational and experimental studies.\citep{Zgarbova2014BaseRNA, Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix} Sanstead et al. highlighted the role of these dynamics during duplex dissociation, where the stability and relaxation timescale of frayed states was dictated by G:C base placement 10-mer oligonucleotides.\citep{Sanstead2016} Together, these dynamics represent substantial deviations from all-or-nothing behavior that can be variably expressed via subtle sequence differences in short oligomers.

In this work, we study the same four sequences explored by Sanstead et al. in an effort to uncover sequence-dependent dynamics and how they relate to metastable structures mentioned above.\citep{Sanstead2016} We use the coarse-grained 3 Sites per Nucleotide (3SPN2) model to perform unbiased simulations of hybridization and dissociation behavior near each sequence's melting temperature.\citep{Hinckley2013AnHybridization} We synthesize the simulation trajectories using Markov State Models (MSMs) to combine many independent and unbiased trajectories and develop an understanding of sequence-specific kinetics and thermodynamics.\citep{Pande2010EverythingAsk} MSMs have recently been implemented to study mechanisms of DNA hybridization, but the slowest sequence-dependent kinetics were not the focus of these studies.\citep{Jin2019, Xiao2019} \citet{Pinamonti2017} used MSMs to compare the slowest dynamics of short RNA oligonucleotides and found that stacking timescales are highly sequence dependent. We take a similar approach to study 10-mer DNA oligonucleotides and implement state free-reversible VAMPnets (SRVs) within the MSM parameterization pipeline to construct higher resolution kinetic models than are achievable using standard time-lagged independent components analysis (tICA) based approaches.

We find that 3SPN2 enables access to long timescales in which we observe sequence-dependent kinetics and relaxation rates that are, within a corrective scaling factor, in agreement with observed experimental trends. Applying SRV-MSMs enables the construction of high-resolution sequence-dependent kinetic models of hybridization and dissociation. Consistent with previous studies, \citep{Phys2014,Romano2013DNADependence,Araque2016LatticeCooperativity} we find the degree of repetitiveness in the sequence -- and therefore the kinetic accessibility and thermodynamic stability of non-native states -- leads to richer dynamics populated by a diversity of long-lived metastable states. Disrupting repetitive stretches of bases by judicious placement of interrupting G:C pairs can tune the landscape from rich 6-state to very simple 2-state behavior. Furthermore, concentrating G:C pairs at the center of a duplex produces a longer-live frayed state that facilitates the hybridization and dissociation process. Analysis within and between metastable state populations using diffusion maps allows us to explore the conformational diversity and heterogeneity within each state at time scales shorter than the kinetic resolution of the SRV-MSMs. This provides us with molecular-level understanding for the differential stability of various conformational states as a function of oligomer sequence. Taken together, our analysis reflects similar results to previous computational and experimental DNA studies, while elucidating new insights into sequence-dependent dynamics, metastable structures, and relative timescales. These results can inform a generalized model between DNA sequence and hybridization mechanism with application towards equilibrium binding technologies such as DNA-PAINT and barcoding.

@@

%Progress has also been made in modeling kinetics for certain dynamic DNA systems, such as toehold exchanges and optical barcodes \citep{Zhang2009ControlExchange, Shah2019ProgrammingFingerprinting}. Sequence effects on association and dissociation rates were predicted using combined experimental and machine learning approaches, however higher order dynamics are less well understood \citep{Zhang2018PredictingSequence}.  Hairpin \citep{Woodside2006DirectAcid} and in duplex hybridization \citep{yer2014KineticsAT-tracts, Sanstead2016, Sanstead2018DirectDehybridization} studies have revealed sequence-dependent effects and varied differential deviations from "all or nothing" behavior. In temperature-jump spectroscopy, G:C placement alters the slow kinetic response attributed to dissociation and a fast response likely corresponding to terminal A:T fraying. Experimentally, however, it is difficult to pick up on dynamics that have similar spectroscopic signatures and to tease apart temporally close responses that might be combined into one stretched exponential signal.

% include Tokmakoff conclusions in the intro -- sequence-dependent deviations from two-state behavior. The slow response was attributed to complete strand dissociation, while the fast response described base fraying. We perform longer equilibrium sampling near Tm to get a sense of sequence equilibrium behavior, as well shorter simulations at a range of higher temperatures to replicate T-jump data. Taken together, our analysis reflects similar results to previous computational and experimental DNA work, while elucidating new insights into sequence-dependent dynamics, metastable structures, and relative timescales.

@@

\subsection{\label{sec:methods}Simulation set up}

The 3SPN2 coarse-grained model was designed to accelerate DNA simulations relative to all-atom models while maintaining experimental melting temperatures, stacking energies, and persistence lengths.\citep{Hinckley2013AnHybridization} Interaction sites are located at phosphate, sugar, and base centers of mass; anisotropic potentials are designed to model non-bonded interactions such as intra-strand base-stacking, inter-strand cross-stacking, and base pairing. The model has been validated against experimentally determined structural properties and hybridization rates, although no dynamic information was used to parameterize the model other than Langevin friction coefficients.\citep{Hinckley2013AnHybridization, Hinckley2014Coarse-grainedEffects} The 3SPN2 model been has widely-adopted to study numerous phenomenon including DNA packing in viral capsids, protein-DNA binding, and nucleosome unwrapping.\citep{Cordoba2017AIons, Lu2020OpenAWSEMSummary, Lequieu2016Tension-dependentUnwrapping}

We initialized four self-complementary double-stranded sequences previously the subject of ultra-fast T-jump experimentation -- 5'ATATATATAT3' (AT-all), 5'GATATATATC3' (GC-end), 5'ATATGCATAT3' (GC-core), and 5'ATGATATCAT3' (GC-mix) -- in a periodic box of dimensions 7.8 nm.\citep{Sanstead2016} We specified a 240 mM implicit salt concentration and used a Debye-Huckel approximation for electrostatic interactions within a 5 nm cutoff. We ran our simulations in the NVT ensemble with a Langevin thermostat that modeled interactions with an implicit water solvent.\citep{Schneider1978Molecular-dynamicsTransitions} Simulations were conducted at empirically determined  melting temperatures in order to observe frequent transitions between dissociated and hybridized states. We used a 20 fs time step and $1.3\times10^{9}$ steps in each simulation, saving frames every 100 ps. For each sequence, 40 simulations were performed in parallel, consuming about 24 serial CPU-hours on 28xIntel E5-2680v4 processors per simulation. Half of runs were initialized from the native hybridized state and half from a dissociated state. The first 10000 frames (1 $\mu$s of simulation time) of each simulation were removed, resulting in 40 x 250000 frames and a total of 1 ms simulation time per sequence. We observed between 55-100 hybridization/dissociation events for each of the four sequences.

\subsection{\label{sec:methods}Featurization}

A featurization of trajectories was required to capture fundamental dynamics while accounting for translational and rotational symmetries. Pairwise distances between base centers of mass -- similar to heavy atom distances in protein models \citep{Sidky}-- satisfied these requirements and were calculated at each frame using the MDtraj Python library.\citep{McGibbon2015MDTraj:Trajectories} Although pairwise distances remove translational and rotational symmetries, additional symmetries arise from the self-complementary nature of these sequences -- the sense and anti-sense strands in each pair are identical to each other. To account for this, we averaged permutable distances (45 pairs in total) together into a "symmetrized" feature set following a similar procedure used in tICAgg coordinate construction.\citep{Sengupta2019AutomatedSelf-assembly} The VAMP-2 scoring method was employed to evaluate the kinetic variance of the feature set and optimize hyperparameters.\citep{WuVariationalData, Mardt2018VAMPnetsKinetics} Additional details on the VAMP-2 score can be found in the supplemental information. By comparing scores across several candidate feature sets (Figure \ref{fig:allseq_features_vamp2}), we found that minimal kinetic information was lost in the  symmetrization procedure. Furthermore, the smaller feature set enabled faster training times and better statistics over permutable distances, without appearing to suffer a loss in generality or model resolution.

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=80mm,
        scale=0.5]{Fig1.pdf}
        \caption{A comparison of VAMP-2 scores via 5-fold cross-validation for the intermolecular, intra+intermolecular,  and symmetrized intermolecular distances, in addition to the reciprocal value for each of these coordinates. Reciprocal coordinates performed better across the board, but there was no substantial difference between the three featurization methods. This indicated that minimal kinetic information was lost in the symmetrization procedure, therefore we chose reciprocal symmetrized coordinates to train the model.}
        \label{fig:allseq_features_vamp2}
	\end{center}
\end{figure}

\subsection{\label{sec:methods}SRVs}

>>>>

%include hyperlinks to figures
%\documentclass{article}
\usepackage{graphicx}
\usepackage[%  
    colorlinks=true,
    pdfborder={0 0 0},
    linkcolor=blue,
    citecolor=black
]{hyperref}

@@

\usepackage[scaled=0.85]{beramono}
\usepackage[T1]{fontenc}
\usepackage{changepage}
\usepackage{mathrsfs}
\usepackage{comment}
\usepackage[normalem]{ulem}
\usepackage{algorithm}
\usepackage{subfig}
\usepackage{algpseudocode}

@@

\newcommand{\state}[1]{$\mathcal{S}_{#1}$}
\newcommand{\Expect}[1]{\mathrm{E}\left[#1\right]}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}

\newcommand*{\rood}[1]{{\color{red}{#1}}}
%\newcommand*{\rood}[1]{#1}
\newcommand*{\blauw}[1]{\textcolor{blue}{#1}}
%\newcommand*{\blauw}[1]{#1}
\newcommand*{\groen}[1]{\textcolor{green}{#1}}
%\newcommand*{\groen}[1]{#1}
%\newcommand*{\blauwr}[1]{{\color{blue}{#1}}}
%\newcommand*{\blauwr}[1]{#1}

\newcommand*{\addref}[1]{\textcolor{red}{\{ADD REF: #1\}}}

@@

%\usepackage[draft]{todonotes}   % notes shown
\usepackage[disable]{todonotes} % notes hidden

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[]{Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy}

@@

\textbf{\Large NOTES}
\begin{itemize}
\item All figs as pdf (makes compilation faster) or tiff
\item For quotation marks in tex use ``XXX'' rather than "XXX"
\item Strong structure is critical

-- It must be a narrative story with a logical flow -- reader must understand where you are taking them next and why

-- Topic sentences

-- Everything is a funnel

\item Figures are included directly after the paragraph after which they are first mentioned

\item We use hybridization / association and dehybridization / dissociation interchangeably. Maybe this is OK, or maybe we should pick one terminology and stick with it.

\item In first mention of anything in SI say, e.g., Fig.~S1 in the Supporting Information. From then on, can just say Fig.~SX.

\item Write out in words numbers less than 10, e.g., one-base, two-times, seven, and use numerals for 10 and greater, e.g. 10-mer, 15-fold, 42 (Only exception I can think of is reporting ranges of numbers, e.g., 2-6.)

\item Every figure caption also starts with a topic sentence. Should also contain, in general, a takeaway message for the reader in addition to a description of the figure content.

\item We probably also need a table of contents (TOC) image. This should be emblematic of the paper as a whole, and cannot be just (any portion of) an existing image. Insert this image at the end of the article after the references.

\end{itemize}

\begin{abstract} % 250 words max for JACS

\noindent A robust understanding of the sequence-dependent thermodynamics of DNA hybridization has enabled rapid advances in DNA nanotechnology. A fundamental understanding of the sequence-dependent kinetics and mechanisms of hybridization and dehybridization remains comparatively underdeveloped. In this work, we establish new understanding of the sequence-dependent hybridization/dehybridization kinetics and mechanism within a family of self-complementary pairs of 10-mer DNA oligomers by integrating coarse-grained molecular simulation, machine learning of the slow dynamical pathways, data-driven inference of long-time kinetic models, and experimental temperature-jump infrared spectroscopy. For a repetitive ATATATATAT sequence, we resolve a rugged dynamical landscape comprising multiple metsatable states, numerous competing hybridization/dehybridization pathways, and a spectrum of dynamical relaxations. Introduction of a G:C pair at the terminus (GATATATATC) or center (ATATGCATAT) of the sequence reduces the ruggedness of the dynamics landscape by eliminating a number of metastable states and reducing the number of competing dynamical pathways. Only by introducing a G:C pair midway between the terminus and the center to maximally disrupt the repetitive nature of the sequence (ATGATATCAT) do we recover a canonical two-state model of hybridization/dehybridization over a smooth landscape with no intermediate metastable states. Our results establish new understanding of the dynamical richness of sequence-dependent kinetics and mechanisms of DNA hybridization/dehybridization, present a molecular basis with which to understand experimental temperature jump data, and furnish foundational design rules by which to rationally engineer the kinetics and pathways of DNA association and dissociation in burgeoning DNA nanotechnology applications.

%\noindent \url{https://www.overleaf.com/project/5e9e5110c524b8000192c548}

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Over the last couple of decades, DNA has proven to be much more than a vessel for genetic information. From sensing to computing to directed self-assembly, the programmable and predictable nature of DNA has unlocked numerous unforeseen nanotechnology applications.\citep{Seeman2017DNANanotechnology, Adleman1994MolecularProblems, Rothemund2006FoldingPatterns, Gu2010ALine} Recently, single molecule localization techniques have exploited the rapid and transient binding of short DNA oligomers in order to achieve super-resolution microscopy and optical multiplexing. \cite{Schnitzbauer2017Super-resolutionDNA-PAINT, Shah2019, Strauss2020UpDNA-PAINT} Predictive understanding of the sequence-dependent thermodynamics of DNA hybridization/dehybridization -- the assembly/disassembly of a DNA duplex from two single strands -- has underpinned the rational design of DNA oligomer sequences for nanotechnological applications, where sequence-dependent nearest-neighbor models can accurately account for mismatched pairs, dangling ends, and other non-native bonding effects.\citep{SantaLucia1998AThermodynamics, Santalucia2004TM} Secondary DNA structures such as hairpins and G-quadruplexes have also been studied in depth and leveraged for nanotechnology applications.\citep{Tsukanov2013DetailedOrigami, Mosayebi2014TheFormation, Mergny2019DNANanotechnology} Predictive models of the \textit{dynamical}, as opposed to purely thermodynamical, behaviors of DNA have become increasingly important in developing technologies such as DNA-PAINT (DNA Points Accumulation For Imaging In Nanoscale Topography), but developments in dynamic DNA nanotechnologies have outpaced our understanding of the dynamics themselves. \citep{Deluca2020DynamicDevices, Cordes2010SensingTransfer, Naimark2020DNADevelopment} Many experimental and computational studies have investigated DNA dynamical phenomena from the picosecond to millisecond range.\citep{Yin2011KineticsHybridization, Xiao2019, Hinckley2014Coarse-grainedEffects, Sanstead2016, Porschke1971CooperativeTransition} Kinetic models have been developed for particular DNA processes such as toehold exchanges and optical barcoding \citep{Zhang2009ControlExchange, Shah2019ProgrammingFingerprinting} \rood{and supervised machine learning techniques have been combined with experimental measurements to predict the on/off rates as a function of sequence.} \citep{Schickinger2018TetheredHelices, Zhang2018PredictingSequence} \noter{Is this red clause correct? Are there simulation components to the work? Was it on/off rates that were predicted?} \rood{A comprehensive understanding of the full dynamical landscape of hybridization/dehybridization accounting for the sequence-dependent metastable states and association/dissociation pathways remains lacking and fundamental questions remain unresolved.} \noter{Accurate? More here? Does this adequately convey wrt to both citations why the two cited investigations are not the final word. i.e., why is there still a need for the present paper -- why is the problem not solved by these works, what did they leave unanswered?} For example, it remains unclear the extent to which hybridization of short DNA oligomers largely proceeds in a conventionally assumed ``all-or-nothing'' fashion or if long-lived metastable states facilitate the transition.\citep{Araque2016LatticeCooperativity, Sikora2013ModelingIntermediates} \noter{Should these references also be included here? \citep{Wyer2014KineticsAT-tracts, Sanstead2016, Sanstead2018DirectDehybridization, Xiao2019}} Out-of-register ``shifted'' base paired structures \citep{Phys2014, Maciejczyk2014DNAModel, Araque2016LatticeCooperativity, Xiao2019} and frayed structures \citep{Zgarbova2014BaseRNA, Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix} stand as candidates for metastable states with the potential to mediate substantial deviations from all-or-nothing behavior, but the extent to which these states are kinetically relevant is difficult to determine experimentally and is likely to be highly sequence-dependent. The development of predictive models and design rules with which to engineer DNA strands with tailored hybridization/dehybridization kinetics and pathways is vital to advancing rational design of DNA strands for nanotechnology applications and is also of importance in understanding fundamental biological processes such as transcription and gene regulation.

Our understanding of hybridization dynamics has been built from decades of experiments -- such as temperature-jump, salt-jump, pH-jump, and other perturbative methods -- that rapidly stimulate DNA and monitor relaxation to a new equilibrium.\citep{Morrison1993SensitiveSolution, Wetmur1968KineticsDNA, Craig1971RelaxationOligonucleotides, Porschke1973ThermodynamicsPairs, Williams1989LaserDGCATGC, Narayanan2012ExploringMixing, Chen2007InfluenceHybridization, Sanstead2018DirectDehybridization} More recently, single molecule diffusion and tethered multifluorophore assays have facilitated equilibrium analysis, but these results can be hampered by slow data collection rates and fluorescent tags effects on strand dynamics, particularly for shorter oligomers.\citep{Liu20173DSolution,  Schickinger2018TetheredHelices, Chen2008Base-by-baseSpectroscopy, Dupuis2013Single-moleculeHelices} A number of computational modeling approaches have also been employed to provide molecular-level resolution of hybridization. Simplified lattice models can recapitulate the essential aspects of the hybridization pathways but lack the realism of continuous space representations \citep{Araque2016LatticeCooperativity, Phys2019}. The long time scales associated with hybridization/dehybridization events place them outside the reach of unbiased all-atom molecular dynamics simulations,\citep{Phys2014} \noter{I moved this citation here -- is that OK?} but they can be observed by employing enhanced sampling techniques such as metadynamics, umbrella sampling,  transition path sampling, and forward flux sampling,\citep{Hinckley2013AnHybridization, Schmitt2013ExploringSurface, Sambriski2009, Sambriski2009SequencePathways, Hoefert2011MolecularOligonucleotides,Romano2013DNADependence, Hinckley2014Coarse-grainedEffects} or by using elevated temperature or denaturing solvent concentrations to induce one-way dissociation events.\citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding} The effect of the applied bias upon the thermodynamics can be rigorously corrected for using standard reweighting techniques, but approaches to rigorously correct the kinetics, particularly under the conditions of high bias necessary for good sampling, are in their infancy. \addref{Girsanov (Keller ? 10.1063/1.4989474, 10.1063/1.5027728 ,10.1137/17M1124772), TRAM (Noe, Hummer ? 10.1073/pnas.1525092113), and MBAR for paths (Chodera, Shirts, Noe ? 10.1063/1.3592152, 10.1063/1.3592153)} A number of coarse-grained DNA force fields have been developed that enable direct observation of these events over microsecond time scales \noter{micro is OK? milli?} via unbiased coarse-grained molecular dynamics (CGMD) simulations, \citep{Romano2013DNADependence, Hinckley2013AnHybridization, Maciejczyk2014DNAModel, Markegard2015, Dans2016MultiscaleDNA} which, up to a speedup factor associated with the smoothing of the underlying free energy landscape inherent to the coarse-graining procedure, can preserve a faithful model of the unbiased dynamics and associated pathways. These models have previously been used to \rood{do X, Y, and Z}. \addref{}

In this work, we study a family of self-complementary pairs of 10-mer DNA oligomers using coarse-grained molecular simulation, machine learning, and data-driven inference of long-time kinetic models to establish new understanding of the influence of sequence upon hybridization/dehybridization kinetics and mechanisms. This family of oligomers -- 5$^\prime$ATATATATAT3$^\prime$ (AT-all), 5$^\prime$GATATATATC3$^\prime$ (GC-end), 5$^\prime$ATATGCATAT3$^\prime$ (GC-core), and 5$^\prime$ATGATATCAT3$^\prime$ (GC-mix) -- was designed to probe the influence of the placement of a G:C pair within an otherwise repetitive A:T sequence and has been the subject of our prior experimental investigations using temperature-jump infrared spectroscopy and simple lattice models. \citep{Sanstead2016} We validate the new computational models of hybridization/dehybridization dynamics developed in this work against new and existing experimental data and reinterpret our prior experimental observations in light of the new computational understanding. Consistent with previous studies, \citep{Phys2014,Romano2013DNADependence,Araque2016LatticeCooperativity} we find the degree of repetitiveness in the sequence -- and therefore the kinetic accessibility and thermodynamic stability of non-native states -- leads to richer dynamics populated by a diversity of long-lived metastable states. Our data-driven modeling and analysis rigorously quantifies these behaviors and furnishes accurate predictive models of the hybridization/dehybridization rates, dynamical pathways, and metastable states. Specifically, we demonstrate that disrupting repetitive stretches of A:T bases by placement of interrupting G:C base pairs enables us to tune the landscape from rich six-state to simple two-state ``all-or-nothing'' behavior, and the specific location of the interrupting pair can be used to modulate the stability of long-lived frayed states that facilitate the hybridization/dehybridization process. Analysis of the configurational ensemble within each metastable state reveals the conformational heterogeneity and dynamical interconversions proceeding on time scales that lie below those resolvable within the long-time kinetic model. Taken together, our analyses establish new molecular-level understanding of the sequence-dependent kinetics and pathways through quantitative predictive models for the long-time system dynamics, resolution of the dynamical folding pathways and metastable states, and elementary design rules with which to sculpt and qualitatively alter the dynamical behaviors of the system. We anticipate that this new foundational understanding, and the extension of our approach to more extensive families of DNA sequences, can guide the rational design of DNA oligomers with tailored kinetic properties engineered for DNA nanotechnology applications such as DNA-PAINT and DNA barcoding. \addref{} \noter{Is barcoding appropriate here, or just PAINT is sufficient?}

@@

% include Tokmakoff conclusions in the intro -- sequence-dependent deviations from two-state behavior. The slow response was attributed to complete strand dissociation, while the fast response described base fraying. We perform longer equilibrium sampling near Tm to get a sense of sequence equilibrium behavior, as well shorter simulations at a range of higher temperatures to replicate T-jump data. Taken together, our analysis reflects similar results to previous computational and experimental DNA work, while elucidating new insights into sequence-dependent dynamics, metastable structures, and relative time scales.

@@

\subsection{Computational Methods}

\subsubsection{Molecular dynamics simulations}

We performed molecular dynamics simulations of four 10-base self-complementary double-stranded DNA sequences that we have previously studied by ultra-fast temperature-jump infrared spectroscopy \citep{Sanstead2016}: 5$^\prime$ATATATATAT3$^\prime$ (AT-all), 5$^\prime$GATATATATC3$^\prime$ (GC-end), 5$^\prime$ATATGCATAT3$^\prime$ (GC-core), and 5$^\prime$ATGATATCAT3$^\prime$ (GC-mix). We modeled the DNA sequences using the coarse-grained 3-Site-Per-Nucleotide v2 (3SPN.2) that uses three spherical beads to represent the phosphate, deoxyribose sugar, and nitrogenous base of each nucleotide and employs anisotropic interaction potentials to accurately treat intra-strand base-stacking, inter-strand cross-stacking, and base pairing. \citep{Hinckley2013AnHybridization} The model was parameterized against experimental data on bond lengths, bend angles, torsional angles, base step energies, and base stacking free energies, and reliably reproduces the structure, melting temperatures, persistence lengths, and sequence, salt, concentration, and temperature effects on duplex formation. \citep{Hinckley2013AnHybridization} The model enables access to millisecond time scales and has been has widely-adopted to study numerous phenomenon including DNA packing in viral capsids, protein-DNA binding, and nucleosome unwrapping.\citep{Cordoba2017AIons, Lu2020OpenAWSEMSummary, Lequieu2016Tension-dependentUnwrapping} Although the 3SPN.2 model was not directly parameterized against dynamical experimental data, we will show below that the predicted sequence-dependent kinetics and relaxation are, within a corrective scaling factor, in very good agreement with observed experimental trends.

All calculations were performed using the LAMMPS simulation package in accordance with best practices for the 3SPN.2 model. \addref{LAMMPS} A single pair of self-complementary sequences were placed in a cubic periodic box with side length 7.8 nm corresponding to a single-strand concentration of \rood{XX} mol/L. Solvent effects were modeled implicitly by employing Langevin dynamics \citep{Schneider1978Molecular-dynamicsTransitions} \noter{Is this the right reference for Langevin dynamics?} with an experimentally motivated per-site friction coefficient of $\xi$ = 9.94 $\times$ 10$^{-11}$ m$^2$/s. \citep{Hinckley2013AnHybridization} \addref{A. E. Nkodo, G. M. Garnier, B. Tinland, C. Desruisseaux, L. C. Mc-Cormick, G. Drouin, and G. W. Slater,Electrophoresis22, 2424 (2001)} We specified a 240 mM implicit salt concentration and treated electrostatic interactions using the Debye-H{\"u}ckel with a 5 nm cutoff radius. \addref{Debye-Huckel} Simulations were performed in the NVT ensemble employing a Langevin thermostat. \addref{Langevin thermostat} Each sequence was simulated at its melting temperature -- AT-all: \rood{XX} K, GC-end: \rood{XX} K, GC-core: \rood{XX} K, GC-mix: \rood{XX} K -- in order to maximize the number of spontaneous transitions between dissociated and hybridized states. Melting temperatures for each sequence were determined empirically by \rood{XXX}. The Langevin equations of motion were integrated using the scheme of Bussi and Parrinello \addref{G. Bussi and M. Parrinello,Phys.Rev.E75, 056707 (2007).} with a 20 fs integration time step. We performed 40 independent simulations for each of the four sequences with half of the runs initialized from the hybridized state and half from the dissociated state. Initial system configurations were prepared \rood{How? How specify initial configurations and apply any energy minimization?} Initial bead velocities assigned from a Maxwell-Boltzmann distribution at the system temperature. \addref{Maxwell-Boltzmann} Each simulation was conducted for 26 $\mu$s and frame saved to disc every 100 ps, requiring $\sim$24 CPU-hours on 28$\times$Intel E5-2680v4 CPU cores. The first 1 $\mu$s of each run was discarded for equilibration providing us with 40$\times$25 $\mu$s = 1 ms of simulation data for each sequence, during which time we observed 55-100 hybridization/dehybridization events.

\subsubsection{Markov state model construction}
